### Introduction

```{r}
title: "workingwithdata"
format: html
```

```{r}
library(tidyverse)
library(haven)
library(descr)
library(haven)
library(labelled)
library(apaThemes)
library(basicstatsbook)
```

### ✅Data Management

At the start of my analysis, I loaded all the necessary libraries. After reviewing the dataset and its coodebook, I selected the variables that would help me answer my research question.

The variables that I selected were **“End of Career Challenges (Physically)”**, ranging from 0–30 (CDCHPHYS), and **“Stopped Dancing for New Career”** (DSNEWCRR). I chose CDCHPHYS because it measures the physical challenges that dancers face at the end of their performance careers, such as injuries or accidents. The DSNEWCRR variable records whether dancers stopped performing to pursue a different career. With these two variables, I am able to address my research question and analyze the relationship between physical challenges and career transitions

### Sort of data

```{r}
file_path<-file.path("..","data","data.sav") 
df<-read_sav(file_path)
head(df)

```

### ✅Frequency Tables

### End of career challenges(Physical)

```{r}
freq(as.ordered(df$CDCHPHYS), plot=FALSE)
```

### Stopped Dancing for New Career

```{r}
freq(as.ordered(df$DSNEWCRR) , plot=FALSE)
```

### **Ethnic Origin**

```{r}
freq(as.ordered(df$A5AETHN1) , plot=FALSE)
```

### Stopped dancing because of health

```{r}
freq(as.ordered(df$WSHEALTH), plot =FALSE)
```

### ✅Univariate Data Visualizations

### Figure 1 Bar Graph

```{r}
#Univariate Graph (categorical variable)
df %>%
  ggplot(aes(x = DSNEWCRR)) + geom_bar()+
  ggtitle("Career Transition")

```

### Bar Graph (Var2)

```{r}
#Univariate Graph (categorical variable)
df %>%
  ggplot(aes(x = A5AETHN1)) + geom_bar()+
  ggtitle("Ethnic Origin")
```

### Bar Graph (3)

```{r}
#Univariate Graph (categorical variable)
df %>%
  ggplot(aes(x =CDCHPHYS )) + geom_bar()+
  ggtitle("End of career")
```

### Career Transition

```{r}
#Univariate Graph (categorical variable)
df %>%
  ggplot(aes(x = DSNEWCRR)) + geom_bar()+
  ggtitle("Career Transition") + 
  xlab("Stop Dancing because of New Career") +
  ylab("Count of People") +
  theme_apa() #from papaja package

```

```{r}
df %>%
  ggplot(aes(x = CDCHPHYS)) +
  geom_bar(binwidth = 5, fill = "lightblue", color = "black") +  # Adds binwidth for better visualization
  ggtitle("End of Career") +
  xlab("End Career") +
  ylab("Count of Individuals") +
  theme_minimal()
```

```{r}
df %>%
  ggplot(aes(x = DSNEWCRR)) +
  geom_bar(binwidth = 5, fill = "lightblue", color = "black") +  # Adds binwidth for better visualization
  ggtitle("Stopped Dancing") +
  xlab("Not dancing") +
  ylab("Count of Individuals") +
  theme_minimal()
```

### ✅Bivariate

### Figure 1

```{r}
# univariate graph ( categorial variable)
df%>%
  filter(!is.na(DSNEWCRR)) %>%
  ggplot(aes(x=DSNEWCRR))+
  geom_bar()+
  ggtitle("career transition") +
  facet_wrap(~CDCHPHYS)

```

```{r}

# Univariate graph (categorical variable)
library(ggplot2)
df %>%
  ggplot(aes(x = CDCHPHYS)) +
  geom_bar(fill = "lightblue", color = "black") +
  ggtitle("Faced Challenges") +
  theme_apa()


```

### Figure 2

```{r}
    
   df%>%
     filter(!is.na(CDCHPHYS)) %>%
  ggplot(aes(x=CDCHPHYS))+
  geom_bar()+
  ggtitle("career transition") +
  facet_wrap(~DSNEWCRR)
    
  
```

```{r}
# Univariate graph (categorical variable)
library(ggplot2)
df %>%
  ggplot(aes(x = DSNEWCRR)) +
  geom_bar(fill = "lightblue", color = "black") +
  ggtitle("Stopped Dancing for New Career") +
  theme_apa()
 
```
